---
title: "Untitleds"
---

$$
L(\alpha, \beta \mid P) = \prod_{i=1}^n \frac{\beta^\alpha P_i^{\alpha-1} e^{-\beta P_i}}{\Gamma(\alpha)}
$$

# Economic Analysis

## Packages

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(readxl)
library(cowplot)
library(gsheet)
library(stats4)
library(minpack.lm)
library(ggdist)
library(lubridate)
library(ggthemes)
library(cowplot)
```

## Import the data

```{r}

eff_yld_data = gsheet2tbl("https://docs.google.com/spreadsheets/d/1kTmVINkk8W_luffQew4nrMvqrvZQhe6JLjR_WzbBQT4/edit?gid=1891034819#gid=1891034819")


cost_data = gsheet2tbl("https://docs.google.com/spreadsheets/d/11K6BssCO1_qMZEb5sU1BMpbNKE0YrACsT4tAY51KvvY/edit?gid=740235580#gid=740235580")

price_data = gsheet2tbl("https://docs.google.com/spreadsheets/d/1dEC3-AckAFJ3XsWTln1KLbz4jOp3d1sV/edit?gid=1164957368#gid=1164957368") %>%
  mutate(data = dmy(data)) |> mutate(year = year(data))
```

## Find the most appropriate theoretical distribution for corn price

```{r}
price_data %>%
  ggplot(aes(preco))+
  geom_histogram(bins = 20,color ="white", fill = "gray")
```

```{r}
price_data %>%
  ggplot(aes(as.factor(year), preco))+
  geom_jitter(width = 0.1)
```

# FIT

How well does the Gamma distribution fit the empirical data?

```{r}
# Empirical cumulative distribution function
environment(ecdf(price_data$preco))

# These are the cumulative values of the ECDF (accumulated probabilities)
Fx = environment(ecdf(price_data$preco))$y
# Original values
x = environment(ecdf(price_data$preco))$x


```

## Nonlinear least squares model

```{r}
# Fits a nonlinear least squares model to the data.
price_reg = nlsLM(Fx ~ pgamma(x, shape, rate, log = FALSE),
      start = c(shape = 2.5, rate = 0.13),
      control = nls.lm.control(maxiter = 1024))
summary(price_reg)

shape = summary(price_reg)$coef[1]
rate = summary(price_reg)$coef[2]

# Plot the ECDF and overlay the Gamma distribution curve
plot(ecdf(price_data$preco))
curve(pgamma(x, shape, rate), 0, 600, add = T, col = "red")
```

## Visualizing the histogram

```{r}
price_data %>%
  ggplot(aes(preco))+
 geom_histogram(aes(y = ..density..),bins = 10, color = "white", fill = "#481567FF")+
   stat_function(fun=function(x) dgamma(x, shape,  rate), size = 1.2, color = "black")+
  theme_bw()+
  xlim(0,20)+
  labs(x="Slope (kg/p.p.)", y = "Density")
```

## **Maximum Likelihood Estimation**

```{r}
MASS::fitdistr(price_data$preco, "gamma", start=list(shape=3.76, rate=0.46))

plot(ecdf(price_data$preco))
curve(pgamma(x, 3.76,  0.46),0,200,  add = T, col = "red")
```

## Visualizing the histogram

```{r}
price_data %>%
  ggplot(aes(preco))+
 geom_histogram(aes(y = ..density..),bins = 15, color = "white", fill = "#481567FF")+
   stat_function(fun=function(x) dgamma(x, 3.76,  0.46), size = 1.2, color = "black")+
  theme_bw()+
  xlim(0,20)+
  labs(x="Slope (kg/p.p.)", y = "Density")
```

## Non linear or MLE

### Kolmogorov-Smirnov Test

The Kolmogorov-Smirnov (KS) test is used to compare two cumulative distributions, testing whether the sample distributions are significantly different. The p-value resulting from the KS test indicates the probability that the data distribution is the same as the theoretical distribution (in this case, the Gamma distribution). If the p-value is small (typically less than 0.05), it suggests that the data distribution differs significantly from the theoretical distribution.

```{r}
# using regression
ks.test(Fx, pgamma(x, shape, rate) )

# using Maximum likelihood
ks.test(Fx, pgamma(x,  3.76,  0.46))

# p-value gives the probability of data distributions be equal to the theoretical distribution obtained 

```

# Probability distributions

## Simulations

```{r}
library(truncnorm)
```

```{r}
set.seed(1)
n_sim = 50000
funs = unique(eff_yld_data$fungicide)
box1 = data.frame()

for(i in 1:length(funs)){
  # Filtering the data for the current fungicide (without using the 'region' variable)
  single_data = eff_yld_data %>% 
    filter(fungicide == funs[i]) # filter for the respective fungicide yield data

  single_cost = cost_data %>% 
    filter(fungicide == funs[i]) # filter for the respective fungicide cost data

  cmin = single_cost$cost - single_cost$cost*0.05 # adjusting a variable cost range for that fungicide
  cmax = single_cost$cost + single_cost$cost*0.05

  C = runif(n_sim, cmin, cmax) # generating n_sim random values (50000) for product price, uniformly distributed

  P = rgamma(n_sim, shape, rate)/60 # generating n_sim random values for corn price based on the equation defined in the previous code

  y_mu = single_data$mean_yld # mean yield for the fungicide
  y_se = single_data$SE_yld # standard error of yield for the fungicide

  Y = rtruncnorm(n_sim, a=0, b=Inf, y_mu, y_se) # ensuring positive values in a truncated distribution

  profit = (P * Y) - C # P(corn price) multiplied by mean yield. Subtract the product cost. Get profit

  eventos = (C <= (P * Y)) * 1 # checking if the cost is less than or equal to profit, if true is 1, otherwise 0.

  # Creating the box for each fungicide
  lil_box = data.frame(fungicide = funs[i],
                       C_point = single_cost$cost, # original cost
                       C, # simulated product cost
                       P,  # simulated corn price
                       Y,   # simulated yield
                       profit, # simulated profits
                       eventos) # less than or equal to profit = 1

  # Adding to the overall box
  box1 = box1 %>% 
    bind_rows(lil_box)
}
```

## Probability of making a profit

```{r}
probs = box1 %>% 
  group_by(fungicide) %>% 
  summarise(P = round(mean(eventos), 2), #probabilidade de sucesso para cada fungicida
            m = mean(profit),
            lower = quantile(profit, 0.025),
            upper = quantile(profit, 0.975)) %>% 
  mutate(P = paste("p =", P))
probs %>% 
  arrange(-m)
```

#Plot

```{r}
box1 |> 
  mutate(fungicide = factor(fungicide, levels = c("TRFX + PROT + BIX","PYRA + FLUX + MEFE",  "AZO + TEBU + MANC", "PYRA + FLUX", "PYRA + EPOX", "DIF + PYDI","MANC", "CLOR"))) |> 
  ggplot(aes()) +
  geom_vline(xintercept = 0, color = "gray40", size = 0.7) +
   stat_slab(aes(profit, fungicide, alpha = stat(plogis(x / 100))),  fill = "#ff9900") +
  #stat_pointinterval(aes(x = profit, y = fungicide), 
   #                  position = position_dodge(width = 0.8, preserve = "single"),  
    #                 .width = c(0.75, 0.95),
     #                size = 1.2, color = "#313695", alpha = 0.8) +
  geom_text(data = probs, 
            aes(x = m, y = fungicide, label = P),
            position = position_dodge(width = 0.8),  
            size = 3, vjust = -3, hjust = -1.5, color = "black", fontface = "bold.italic") +  
  geom_point(data = cost_data,
             aes(spray, fungicide, shape = "Spray cost"),
             size =2, color = "#d73027" )+
  geom_text(data = cost_data,
          aes(x = spray, y = fungicide, label = sprintf("%.1f", spray)),
          size = 3, color = "#d73027", vjust = -0.5, hjust= 0, fontface = "bold")+
  guides(alpha = "none", shape = "none", fill = guide_legend(title = "Profit Status")) +
 theme_minimal_vgrid() +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        panel.grid.major.y = element_blank(), 
        panel.grid.minor = element_blank(),
         panel.background = element_rect(fill = scales::alpha("white", 0.01)),
        plot.margin = unit(c(1, 1, 1, 1), "cm")) +
  coord_cartesian(xlim = c(-200, 450)) +
   scale_x_continuous(breaks = seq(0, 450, by = 50))+
  labs(x = expression(bold("Profit (US$ ha"^-1*")")),
     y = expression(bold("Fungicide")),
       fill = "Profit Status",
       color = "Profit Status",
       shape = "Cost")+
 geom_point(data = probs, aes(x = m, y = fungicide), size = 2, color = "#313695", position = position_nudge(y = -0.2)) +
  geom_errorbarh(data = probs, aes(xmin = lower, xmax = upper, y = fungicide), 
                 height = 0, size= 0.6, color = "#313695", position = position_nudge(y = -0.2))+
geom_text(data = probs, 
          aes(x = m, y = fungicide, label = paste0("π = ", round(m, 1))),
          position = position_dodge(width = 0.8),  
          size = 3, vjust = -1.5, hjust = -1.5, color = "#313695", fontface = "bold.italic")
library(patchwork)

#ggsave("figures/profit.png", dpi=600, height = 7,bg = "white", width = 7)
```
